2-factor-Hull-White\_model\_calibration

### 2020-05-12 week7/Report.tex
Report Compiled

\section{LIBOR forward rates, stock paths, and counterparty firm values.
}

\section{Discount factor and Value of the Option
}

\section{Conclusion}

>>>>

\section{LIBOR Forward Rates, Stock Paths, and Counterparty Firm Values
}

In this part, we use a sample size of 100000, jointly simulate LIBOR forward rates, stock paths, and counterparty firm values. We assume that the counterparty firm and stock values are uncorrelated with LIBOR forward rates.

\subsection{Calibrate LIBOR forward rate model from zero coupon bond prices
}

    We initialize the given zero-coupon bond prices:

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{4}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{t} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{linspace}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+m+mi}{13}\PY{p}{)}

\PY{n}{market\PYZus{}zcb\PYZus{}prices} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{p}{[}\PY{l+m+mf}{1.0}\PY{p}{,} \PY{l+m+mf}{0.9938}\PY{p}{,} \PY{l+m+mf}{0.9876}\PY{p}{,} \PY{l+m+mf}{0.9815}\PY{p}{,} \PY{l+m+mf}{0.9754}\PY{p}{,} \PY{l+m+mf}{0.9694}\PY{p}{,} \PY{l+m+mf}{0.9634}\PY{p}{,} \PY{l+m+mf}{0.9574}\PY{p}{,} \PY{l+m+mf}{0.9516}\PY{p}{,}
       \PY{l+m+mf}{0.9457}\PY{p}{,} \PY{l+m+mf}{0.9399}\PY{p}{,} \PY{l+m+mf}{0.9342}\PY{p}{,} \PY{l+m+mf}{0.9285}\PY{p}{]}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    We next create functions to calculate the simulated bond prices from the
Vasicek model (as well as helper functions A and D). We also define
function F which is the differences between the bond prices calculated
by our model and actual market zero-coupon bond prices:

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{5}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{def} \PY{n+nf}{A}\PY{p}{(}\PY{n}{t1}\PY{p}{,} \PY{n}{t2}\PY{p}{,} \PY{n}{alpha}\PY{p}{)}\PY{p}{:}
    \PY{k}{return} \PY{p}{(}\PY{l+m+mi}{1}\PY{o}{\PYZhy{}}\PY{n}{np}\PY{o}{.}\PY{n}{exp}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{n}{alpha}\PY{o}{*}\PY{p}{(}\PY{n}{t2}\PY{o}{\PYZhy{}}\PY{n}{t1}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{o}{/}\PY{n}{alpha}
\PY{k}{def} \PY{n+nf}{D}\PY{p}{(}\PY{n}{t1}\PY{p}{,} \PY{n}{t2}\PY{p}{,} \PY{n}{alpha}\PY{p}{,} \PY{n}{b}\PY{p}{,} \PY{n}{sigma}\PY{p}{)}\PY{p}{:}
    \PY{n}{val1} \PY{o}{=} \PY{p}{(}\PY{n}{t2}\PY{o}{\PYZhy{}}\PY{n}{t1}\PY{o}{\PYZhy{}}\PY{n}{A}\PY{p}{(}\PY{n}{t1}\PY{p}{,}\PY{n}{t2}\PY{p}{,}\PY{n}{alpha}\PY{p}{)}\PY{p}{)}\PY{o}{*}\PY{p}{(}\PY{n}{sigma}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{o}{/}\PY{p}{(}\PY{l+m+mi}{2}\PY{o}{*}\PY{n}{alpha}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{p}{)}\PY{o}{\PYZhy{}}\PY{n}{b}\PY{p}{)}
    \PY{n}{val2} \PY{o}{=} \PY{n}{sigma}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{o}{*}\PY{n}{A}\PY{p}{(}\PY{n}{t1}\PY{p}{,}\PY{n}{t2}\PY{p}{,}\PY{n}{alpha}\PY{p}{)}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{o}{/}\PY{p}{(}\PY{l+m+mi}{4}\PY{o}{*}\PY{n}{alpha}\PY{p}{)}
    \PY{k}{return} \PY{n}{val1}\PY{o}{\PYZhy{}}\PY{n}{val2}

\PY{k}{def} \PY{n+nf}{bond\PYZus{}price\PYZus{}fun}\PY{p}{(}\PY{n}{r}\PY{p}{,}\PY{n}{t}\PY{p}{,}\PY{n}{T}\PY{p}{,} \PY{n}{alpha}\PY{p}{,} \PY{n}{b}\PY{p}{,} \PY{n}{sigma}\PY{p}{)}\PY{p}{:}
    \PY{k}{return} \PY{n}{np}\PY{o}{.}\PY{n}{exp}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{n}{A}\PY{p}{(}\PY{n}{t}\PY{p}{,}\PY{n}{T}\PY{p}{,}\PY{n}{alpha}\PY{p}{)}\PY{o}{*}\PY{n}{r}\PY{o}{+}\PY{n}{D}\PY{p}{(}\PY{n}{t}\PY{p}{,}\PY{n}{T}\PY{p}{,}\PY{n}{alpha}\PY{p}{,}\PY{n}{b}\PY{p}{,}\PY{n}{sigma}\PY{p}{)}\PY{p}{)}

\PY{k}{def} \PY{n+nf}{F}\PY{p}{(}\PY{n}{x}\PY{p}{)}\PY{p}{:}
    \PY{n}{alpha} \PY{o}{=} \PY{n}{x}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}
    \PY{n}{b} \PY{o}{=} \PY{n}{x}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}
    \PY{n}{sigma} \PY{o}{=} \PY{n}{x}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}
    \PY{n}{r0} \PY{o}{=} \PY{n}{x}\PY{p}{[}\PY{l+m+mi}{3}\PY{p}{]}
    \PY{k}{return} \PY{n+nb}{sum}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{abs}\PY{p}{(}\PY{n}{bond\PYZus{}price\PYZus{}fun}\PY{p}{(}\PY{n}{r0}\PY{p}{,}\PY{l+m+mi}{0}\PY{p}{,}\PY{n}{t}\PY{p}{,}\PY{n}{alpha}\PY{p}{,}\PY{n}{b}\PY{p}{,}\PY{n}{sigma}\PY{p}{)}\PY{o}{\PYZhy{}}\PY{n}{market\PYZus{}zcb\PYZus{}prices}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    We use the fmin\_slsqp function from scipy to calculate the optimal
model parameters, with a minimum value of F

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{6}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{}minimizing F}
\PY{n}{bnds} \PY{o}{=} \PY{p}{(}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+m+mf}{0.2}\PY{p}{)}\PY{p}{,}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+m+mf}{0.2}\PY{p}{)}\PY{p}{,} \PY{p}{(}\PY{l+m+mf}{0.00}\PY{p}{,}\PY{l+m+mf}{0.10}\PY{p}{)}\PY{p}{)}
\PY{n}{opt\PYZus{}val} \PY{o}{=} \PY{n}{scipy}\PY{o}{.}\PY{n}{optimize}\PY{o}{.}\PY{n}{fmin\PYZus{}slsqp}\PY{p}{(}\PY{n}{F}\PY{p}{,} \PY{p}{(}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{l+m+mf}{0.05}\PY{p}{,} \PY{l+m+mf}{0.03}\PY{p}{,} \PY{l+m+mf}{0.05}\PY{p}{)}\PY{p}{,} \PY{n}{bounds}\PY{o}{=}\PY{n}{bnds}\PY{p}{)}
\PY{n}{opt\PYZus{}alpha} \PY{o}{=} \PY{n}{opt\PYZus{}val}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}
\PY{n}{opt\PYZus{}b} \PY{o}{=} \PY{n}{opt\PYZus{}val}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}
\PY{n}{opt\PYZus{}sigma} \PY{o}{=} \PY{n}{opt\PYZus{}val}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}
\PY{n}{opt\PYZus{}r0} \PY{o}{=} \PY{n}{opt\PYZus{}val}\PY{p}{[}\PY{l+m+mi}{3}\PY{p}{]}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
Optimization terminated successfully.    (Exit mode 0)
            Current function value: 0.0002564991813429618
            Iterations: 10
            Function evaluations: 74
            Gradient evaluations: 10
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{7}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Optimal alpha: }\PY{l+s+si}{\PYZob{}:.3f\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{opt\PYZus{}val}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{)}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Optimal b: }\PY{l+s+si}{\PYZob{}:.3f\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{opt\PYZus{}val}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Optimal sigma }\PY{l+s+si}{\PYZob{}:.3f\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{opt\PYZus{}val}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{p}{)}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Optimal r0: }\PY{l+s+si}{\PYZob{}:.3f\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{opt\PYZus{}val}\PY{p}{[}\PY{l+m+mi}{3}\PY{p}{]}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
Optimal alpha: 0.273
Optimal b: 0.069
Optimal sigma 0.028
Optimal r0: 0.075
    \end{Verbatim}

    We plot the actual market bond prices, with model-derived bond prices,
and they look like a close fit.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{8}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{model\PYZus{}prices} \PY{o}{=} \PY{n}{bond\PYZus{}price\PYZus{}fun}\PY{p}{(}\PY{n}{opt\PYZus{}r0}\PY{p}{,}\PY{l+m+mi}{0}\PY{p}{,}\PY{n}{t}\PY{p}{,} \PY{n}{opt\PYZus{}alpha}\PY{p}{,} \PY{n}{opt\PYZus{}b}\PY{p}{,} \PY{n}{opt\PYZus{}sigma}\PY{p}{)}
\PY{n}{model\PYZus{}yield} \PY{o}{=} \PY{o}{\PYZhy{}}\PY{n}{np}\PY{o}{.}\PY{n}{log}\PY{p}{(}\PY{n}{model\PYZus{}prices}\PY{p}{)}\PY{o}{/}\PY{n}{t}

\PY{n}{plt}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{t}\PY{p}{,}\PY{n}{market\PYZus{}zcb\PYZus{}prices}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Market Prices}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{plt}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{t}\PY{p}{,} \PY{n}{model\PYZus{}prices}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{.}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Calibrated Prices}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{p}{)}
\PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{images/output_15_1.png}
    \end{center}
    { \hspace*{\fill} \\}

\subsection{Simulate LIBOR rate paths
}

    We first initialize the parameter \(\sigma_j\)

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{9}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{sigmaj} \PY{o}{=} \PY{l+m+mf}{0.2}
\end{Verbatim}
\end{tcolorbox}

    We use paramters we obtained above to recreate the Vasicek bond prices:

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{10}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{def} \PY{n+nf}{A}\PY{p}{(}\PY{n}{t1}\PY{p}{,} \PY{n}{t2}\PY{p}{)}\PY{p}{:}
    \PY{k}{return} \PY{p}{(}\PY{l+m+mi}{1}\PY{o}{\PYZhy{}}\PY{n}{np}\PY{o}{.}\PY{n}{exp}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{n}{opt\PYZus{}alpha}\PY{o}{*}\PY{p}{(}\PY{n}{t2}\PY{o}{\PYZhy{}}\PY{n}{t1}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{o}{/}\PY{n}{opt\PYZus{}alpha}

\PY{k}{def} \PY{n+nf}{C}\PY{p}{(}\PY{n}{t1}\PY{p}{,} \PY{n}{t2}\PY{p}{)}\PY{p}{:}
    \PY{n}{val1} \PY{o}{=} \PY{p}{(}\PY{n}{t2}\PY{o}{\PYZhy{}}\PY{n}{t1}\PY{o}{\PYZhy{}}\PY{n}{A}\PY{p}{(}\PY{n}{t1}\PY{p}{,}\PY{n}{t2}\PY{p}{)}\PY{p}{)}\PY{o}{*}\PY{p}{(}\PY{n}{opt\PYZus{}sigma}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{o}{/}\PY{p}{(}\PY{l+m+mi}{2}\PY{o}{*}\PY{n}{opt\PYZus{}alpha}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{p}{)}\PY{o}{\PYZhy{}}\PY{n}{opt\PYZus{}b}\PY{p}{)}
    \PY{n}{val2} \PY{o}{=} \PY{n}{opt\PYZus{}sigma}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{o}{*}\PY{n}{A}\PY{p}{(}\PY{n}{t1}\PY{p}{,}\PY{n}{t2}\PY{p}{)}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{o}{/}\PY{p}{(}\PY{l+m+mi}{4}\PY{o}{*}\PY{n}{opt\PYZus{}alpha}\PY{p}{)}
    \PY{k}{return} \PY{n}{val1} \PY{o}{\PYZhy{}} \PY{n}{val2}

\PY{k}{def} \PY{n+nf}{bond\PYZus{}price}\PY{p}{(}\PY{n}{r}\PY{p}{,}\PY{n}{t}\PY{p}{,}\PY{n}{T}\PY{p}{)}\PY{p}{:}
    \PY{k}{return} \PY{n}{np}\PY{o}{.}\PY{n}{exp}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{n}{A}\PY{p}{(}\PY{n}{t}\PY{p}{,}\PY{n}{T}\PY{p}{)}\PY{o}{*}\PY{n}{r}\PY{o}{+}\PY{n}{C}\PY{p}{(}\PY{n}{t}\PY{p}{,}\PY{n}{T}\PY{p}{)}\PY{p}{)}

\PY{n}{vasi\PYZus{}bond} \PY{o}{=} \PY{n}{bond\PYZus{}price}\PY{p}{(}\PY{n}{opt\PYZus{}r0}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{n}{t}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    The prices calculated from the Vasicek model are close to the ZCB prices
given by the assignment:

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{11}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n+nb}{print}\PY{p}{(}\PY{n}{vasi\PYZus{}bond}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
[1.         0.99377572 0.98760087 0.98147516 0.97539831 0.96936998
 0.96338983 0.95745752 0.95157266 0.94573489 0.93994381 0.93419901
 0.9285001 ]
    \end{Verbatim}

    We now initialize the matrices we will use to store the Monte Carlo
simulations, for both basic Monte Carlo and Predictor-Corrector method.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{12}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{n\PYZus{}simulations} \PY{o}{=} \PY{n}{sample\PYZus{}size}
\PY{n}{n\PYZus{}steps} \PY{o}{=} \PY{n+nb}{len}\PY{p}{(}\PY{n}{t}\PY{p}{)}

\PY{n}{mc\PYZus{}forward} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{ones}\PY{p}{(}\PY{p}{[}\PY{n}{n\PYZus{}simulations}\PY{p}{,} \PY{n}{n\PYZus{}steps}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}\PY{o}{*}\PY{p}{(}\PY{n}{vasi\PYZus{}bond}\PY{p}{[}\PY{p}{:}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{\PYZhy{}}\PY{n}{vasi\PYZus{}bond}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{]}\PY{p}{)}\PY{o}{/}\PY{p}{(}\PY{n}{vasi\PYZus{}bond}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{]}\PY{p}{)}
\PY{n}{predcorr\PYZus{}forward} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{ones}\PY{p}{(}\PY{p}{[}\PY{n}{n\PYZus{}simulations}\PY{p}{,} \PY{n}{n\PYZus{}steps}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}\PY{o}{*}\PY{p}{(}\PY{n}{vasi\PYZus{}bond}\PY{p}{[}\PY{p}{:}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{\PYZhy{}}\PY{n}{vasi\PYZus{}bond}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{]}\PY{p}{)}\PY{o}{/}\PY{p}{(}\PY{n}{vasi\PYZus{}bond}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{]}\PY{p}{)}
\PY{n}{predcorr\PYZus{}capfac} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{ones}\PY{p}{(}\PY{p}{[}\PY{n}{n\PYZus{}simulations}\PY{p}{,} \PY{n}{n\PYZus{}steps}\PY{p}{]}\PY{p}{)}
\PY{n}{mc\PYZus{}capfac} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{ones}\PY{p}{(}\PY{p}{[}\PY{n}{n\PYZus{}simulations}\PY{p}{,} \PY{n}{n\PYZus{}steps}\PY{p}{]}\PY{p}{)}

\PY{n}{delta} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{ones}\PY{p}{(}\PY{p}{[}\PY{n}{n\PYZus{}simulations}\PY{p}{,} \PY{n}{n\PYZus{}steps} \PY{o}{\PYZhy{}} \PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}\PY{o}{*}\PY{p}{(}\PY{n}{t}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{]}\PY{o}{\PYZhy{}}\PY{n}{t}\PY{p}{[}\PY{p}{:}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    We now run the Monte Carlo simulation for each time step:

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{13}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{n\PYZus{}steps}\PY{p}{)}\PY{p}{:}
    \PY{n}{Z} \PY{o}{=} \PY{n}{norm}\PY{o}{.}\PY{n}{rvs}\PY{p}{(}\PY{n}{size}\PY{o}{=}\PY{p}{[}\PY{n}{n\PYZus{}simulations}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}

    \PY{n}{muhat} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{cumsum}\PY{p}{(}\PY{n}{delta}\PY{p}{[}\PY{p}{:}\PY{p}{,} \PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{o}{*}\PY{n}{mc\PYZus{}forward}\PY{p}{[}\PY{p}{:}\PY{p}{,} \PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{o}{*}\PY{n}{sigmaj}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{o}{/}\PY{p}{(}\PY{l+m+mi}{1}\PY{o}{+}\PY{n}{delta}\PY{p}{[}\PY{p}{:}\PY{p}{,} \PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{o}{*}\PY{n}{mc\PYZus{}forward}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{p}{)}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
    \PY{n}{mc\PYZus{}forward}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]} \PY{o}{=} \PY{n}{mc\PYZus{}forward}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{exp}\PY{p}{(}\PY{p}{(}\PY{n}{muhat}\PY{o}{\PYZhy{}}\PY{n}{sigmaj}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{o}{/}\PY{l+m+mi}{2}\PY{p}{)}\PY{o}{*}\PY{n}{delta}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{o}{+}\PY{n}{sigmaj}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{sqrt}\PY{p}{(}\PY{n}{delta}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{p}{)}\PY{o}{*}\PY{n}{Z}\PY{p}{)}

    \PY{n}{mu\PYZus{}initial} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{cumsum}\PY{p}{(}\PY{n}{delta}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{o}{*}\PY{n}{predcorr\PYZus{}forward}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{o}{*}\PY{n}{sigmaj}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{o}{/}\PY{p}{(}\PY{l+m+mi}{1}\PY{o}{+}\PY{n}{delta}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{o}{*}\PY{n}{predcorr\PYZus{}forward}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{p}{)}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
    \PY{n}{for\PYZus{}temp} \PY{o}{=} \PY{n}{predcorr\PYZus{}forward}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{exp}\PY{p}{(}\PY{p}{(}\PY{n}{mu\PYZus{}initial}\PY{o}{\PYZhy{}}\PY{n}{sigmaj}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{o}{/}\PY{l+m+mi}{2}\PY{p}{)}\PY{o}{*}\PY{n}{delta}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{o}{+}\PY{n}{sigmaj}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{sqrt}\PY{p}{(}\PY{n}{delta}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{p}{)}\PY{o}{*}\PY{n}{Z}\PY{p}{)}
    \PY{n}{mu\PYZus{}term} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{cumsum}\PY{p}{(}\PY{n}{delta}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{o}{*}\PY{n}{for\PYZus{}temp}\PY{o}{*}\PY{n}{sigmaj}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{o}{/}\PY{p}{(}\PY{l+m+mi}{1}\PY{o}{+}\PY{n}{delta}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{o}{*}\PY{n}{for\PYZus{}temp}\PY{p}{)}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
    \PY{n}{predcorr\PYZus{}forward}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]} \PY{o}{=} \PY{n}{predcorr\PYZus{}forward}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{exp}\PY{p}{(}\PY{p}{(}\PY{n}{mu\PYZus{}initial}\PY{o}{+}\PY{n}{mu\PYZus{}term}\PY{o}{\PYZhy{}}\PY{n}{sigmaj}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{p}{)}\PY{o}{*}\PY{n}{delta}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{o}{/}\PY{l+m+mi}{2}\PY{o}{+}\PY{n}{sigmaj}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{sqrt}\PY{p}{(}\PY{n}{delta}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{i}\PY{p}{:}\PY{p}{]}\PY{p}{)}\PY{o}{*}\PY{n}{Z}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    From our Monte Carlo simulation, we now calculate the capitalization
factors and bond prices, and plot them to compare them with the Vasicek
bond prices.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{14}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{mc\PYZus{}capfac}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{]} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{cumprod}\PY{p}{(}\PY{l+m+mi}{1}\PY{o}{+}\PY{n}{mc\PYZus{}forward}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
\PY{n}{predcorr\PYZus{}capfac}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{]} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{cumprod}\PY{p}{(}\PY{l+m+mi}{1}\PY{o}{+}\PY{n}{predcorr\PYZus{}forward}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}

\PY{n}{mc\PYZus{}price} \PY{o}{=} \PY{n}{mc\PYZus{}capfac}\PY{o}{*}\PY{o}{*}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}
\PY{n}{predcorr\PYZus{}price} \PY{o}{=} \PY{n}{predcorr\PYZus{}capfac}\PY{o}{*}\PY{o}{*}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}
